---
title: "BDA3 Exercises: 1"
author: "Oscar Oelrich"
date: "11 november 2018"
output: github_document
---



Suggested solutions to some of the exercises in Bayesian Data Analysis 3rd edition by Gelman et al.

# Chapter 1

## Exercise 1.1

### Part a

Since $\theta$ is a discrete variable which takes only the values $1$ and $2$, the marginal probability density for $y$ will be obtained by summing over $\theta$

$$\begin{equation}
p(y) = \sum_{i=1}^2 p(y|\theta_i) p(\theta_i) = \frac{1}{2}p(y|\theta=1)+\frac{1}{2}p(y|\theta=2)
\end{equation}$$
. Using $\sigma=2$ the density will be given by

$$\begin{equation}
p(y)=\frac{1}{2} \frac{1}{\sqrt{2\pi 2^2}}e^{-(y-1)^2/(2*2^2)}   + \frac{1}{2}\frac{1}{\sqrt{2\pi 2^2}}e^{-(y-2)^2/(2*2^2)}= \frac{1}{4\sqrt{2\pi}}\left(e^{-(y-1)^2/(2*2^2)}+e^{-(y-2)^2/(2*2^2)}\right)
\end{equation}$$

We plot the density

```{r}
p1 <- function(y) 0.25*(1/sqrt(2*pi))*(exp(-(y-1)^2/(2*2^2))+exp(-(y-2)^2/(2*2^2)))
# Or we could just use the build in densities from r...
#p2 <- function(y) 0.5*dnorm(y, 1, 2)+0.5*dnorm(y,2,2)
x <- seq(-5, 8, 0.001)
plot(x, p1(x), type = 'l')
```



### Part b

Using Bayes' theorem we have that

$$\begin{equation}
p(\theta=1|y=1) = \frac{p(y=1|\theta=1) p(\theta=1)}{p(y=1)}
\end{equation}$$

which we calculate to be

```{r}
dnorm(1,1,2)*0.5/(0.5*dnorm(1, 1, 2)+0.5*dnorm(1,2,2))
```

That is, after having observed $y=1$ we update our believes favouring $\theta=1$, but not by much.

### Part c


Given that we have observed $y=1$, decreasing the value of the variance means that the observation becomes less and less probable under the alternative $\theta=2$ and so the posterior will put more and more weight on $\theta=1$ as the variance decreases. As the variance increases, the probabilities under the two values will creep closer to each other and the posterior will get closer to the prior of $\frac{1}{2}$.

## Exercise 1.2

We wish to show that the following hold when $u$ is a vector

$$\begin{equation}
\mathbb{E}(u) = \mathbb{E}(\mathbb{E}(u|v))
\end{equation}$$

$$\begin{equation}
\text{var}(u) = \mathbb{E}(\text{var}(u|v))+\text{var}(\mathbb{E}(u|v))
\end{equation}$$

We start with the conditional mean. 


## Exercise 1.6

Approximately $1/125$ of all births are fraternal twins, and approximately $1/300$ of all births are identical twins. Elvis Presley had a twin brother who died at birth. Given the approximation that half of all births are boys, what is the probability that Elvis had a twin brother?

We will solve the problem using Bayes' theorem. We use the following notation.

* *IT* is the event of an identical twin
* *FT* is the event of a fraternal twin
* *T* denotes the probability of having a twin and is obtained by summing $IT$ and $FT$
* *B* denotes the probability of having a brother


We then have that

$$\begin{equation}
P((IT)|(T,B))=\frac{P(T,B|(IT))P(IT)}{P(T,B)}
\end{equation}$$

where we have that $P(T,B|IT)=1$, since if Elvis has an identical twin, obviously it's a brother, the probability for $IT$ is given and the probability $P(T,B)=P(IT,B)+P(FT,B)=P(IT)+P(B|FT)P(FT)=1/300+1/2 *1/125$ which gives the solution

$$\begin{equation}
P((IT)|(T,B))=\frac{1*\frac{1}{300}}{\frac{1}{300}+\frac{1}{2}\frac{1}{125}}=\frac{5}{11}
\end{equation}$$




