---
title: 'BDA3 Exercises: 1'
author: "Oscar Oelrich"
date: "11 november 2018"
output:
  html_document:
    df_print: paged
---



Suggested solutions to some of the exercises in Bayesian Data Analysis 3rd edition by Gelman et al.

# Chapter 1

## Exercise 1.1

### Part a

Since $\theta$ is a discrete variable which takes only the values $1$ and $2$, the marginal probability density for $y$ will be obtained by summing over $\theta$

$$\begin{equation}
p(y) = \sum_{i=1}^2 p(y|\theta_i) p(\theta_i) = \frac{1}{2}p(y|\theta=1)+\frac{1}{2}p(y|\theta=2)
\end{equation}$$
. Using $\sigma=2$ the density will be given by

$$\begin{equation}
p(y)=\frac{1}{2} \frac{1}{\sqrt{2\pi 2^2}}e^{-(y-1)^2/(2*2^2)}   + \frac{1}{2}\frac{1}{\sqrt{2\pi 2^2}}e^{-(y-2)^2/(2*2^2)}= \frac{1}{4\sqrt{2\pi}}\left(e^{-(y-1)^2/(2*2^2)}+e^{-(y-2)^2/(2*2^2)}\right)
\end{equation}$$

We plot the density

```{r}
p1 <- function(y) 0.25*(1/sqrt(2*pi))*(exp(-(y-1)^2/(2*2^2))+exp(-(y-2)^2/(2*2^2)))
# Or we could just use the build in densities from r...
#p2 <- function(y) 0.5*dnorm(y, 1, 2)+0.5*dnorm(y,2,2)
x <- seq(-5, 8, 0.001)
plot(x, p1(x), type = 'l')
```



### Part b

Using Bayes' theorem we have that

$$\begin{equation}
p(\theta=1|y=1) = \frac{p(y=1|\theta=1) p(\theta=1)}{p(y=1)}
\end{equation}$$

which we calculate to be

```{r}
dnorm(1,1,2)*0.5/(0.5*dnorm(1, 1, 2)+0.5*dnorm(1,2,2))
```

That is, after having observed $y=1$ we update our believes favouring $\theta=1$, but not by much.

### Part c


Given that we have observed $y=1$, decreasing the value of the variance means that the observation becomes less and less probable under the alternative $\theta=2$ and so the posterior will put more and more weight on $\theta=1$ as the variance decreases. As the variance increases, the probabilities under the two values will creep closer to each other and the posterior will get closer to the prior of $\frac{1}{2}$.

## Exercise 1.2

We wish to show that the following hold when $u$ is a vector

$$\begin{equation}
\mathbb{E}(u) = \mathbb{E}(\mathbb{E}(u|v))
\end{equation}$$

$$\begin{equation}
\text{var}(u) = \mathbb{E}(\text{var}(u|v))+\text{var}(\mathbb{E}(u|v))
\end{equation}$$

We start with the conditional mean. 


## Exercise 1.6

Approximately $1/125$ of all births are fraternal twins, and approximately $1/300$ of all births are identical twins. Elvis Presley had a twin brother who died at birth. Given the approximation that half of all births are boys, what is the probability that Elvis had a twin brother?

We will solve the problem using Bayes' theorem. We use the following notation.

* *IT* is the event of an identical twin
* *FT* is the event of a fraternal twin
* *T* denotes the probability of having a twin and is obtained by summing $IT$ and $FT$
* *B* denotes the probability of having a brother


We then have that

$$\begin{equation}
P((IT)|(T,B))=\frac{P(T,B|(IT))P(IT)}{P(T,B)}
\end{equation}$$

where we have that $P(T,B|IT)=1$, since if Elvis has an identical twin, obviously it's a brother, the probability for $IT$ is given and the probability $P(T,B)=P(IT,B)+P(FT,B)=P(IT)+P(B|FT)P(FT)=1/300+1/2 *1/125$ which gives the solution

$$\begin{equation}
P((IT)|(T,B))=\frac{1*\frac{1}{300}}{\frac{1}{300}+\frac{1}{2}\frac{1}{125}}=\frac{5}{11}
\end{equation}$$

## Exerecise 1.7

In this exercise we will use Bayes theorem to solve the Monty Hall problem. The key to solving it is in how we chose to condition. Without loss of generality we call the door we choose **A**, the door Monty opens **B** and the last door (that we can potentially switch to) door **C**. We want to calculate the probability of winning when using the two different strategies available to us (switching and not switching). Since we picked door **A**, the probability we are interested in when we want to find the probability of winning using the "not switch" strategy is the probability that the car is behind door **A** given that Monty opened door **B**. We use $M(-)$ to denote Monty opening door -, and $C(-)$ for the car being behind door -.


$$\begin{equation}
P(C(A)|M(B))=\frac{P(M(B)|C(A))P(C(A))}{P(M(B))}
\end{equation}$$

To calculate this probability we need to calculate the following.

$$\begin{itemize}
\item $P(M(B)|C(A))$, which is the probability that Monty opens door **B** given that the car is behind door **A**. Since the car is behind **A**, Monty can chose between door **B** and **C** freely, and so this probability is $0.5$.
\item The unconditional probability that the car is behind any particular door is $P(C(C))=\frac{1}{3}$
\item The unconditional probability that Monty opens door **B** can be obtained by summing over the conditional probabilities for the car being behind the different doors. If the car is being door **B** it is zero, if the car is behind door **C** it is one and if the car is behind door **A** it is one half, with the probability of the car being behind any particular door being one third for each door, this sums up to one $\frac{1}{2}$.
\end{itemize}$$

Plugging in the values into Bayes theorem we obtain

$$\begin{equation}
P(C(A)|M(B))=\frac{\frac{1}{2}\frac{1}{3}}{\frac{1}{2}}=\frac{1}{3}
\end{equation}$$

If we instead use the "switching" strategy, we only need to calculate one new probability, namely $P(M(B)|C(C))$, since now we are interested in $P(C(C)|M(B))$ (which is the probability of winning by using the switching strategy, we win if the car is behind the door we did not chose and that Monty did not open). Given that the car is behind door **C**, Monty has no choice but to open door **B** so this probability is one. This gives us a chance of winning equal to

$$\begin{equation}
P(C(C)|M(B))=\frac{1\frac{1}{3}}{\frac{1}{2}}=\frac{2}{3}
\end{equation}$$

Note that we could have obtained this by symmetry. Since given that Monty opens door **B**, the car is either behind door **A** or **C**. So if we calculate the probability of it being behind **A** to be one third, then the probability that it is behind door **C** is clearly two thirds.



