---
title: "Exercise 1.9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


New patients arrive at a doctors office between 9 in the morning and 4 in the afternoon. The time between the arrival of patients follows an exponential distribution with expectation 10 minutes.

Each patient requires between 5 and 20 minutes of doctor time.

### Part a


We want to simulate the number of patients that arrive at the office during one day. We simulate from an exponential distribution with mean 10 until we hit a total of 420 minutes (at which time the doctor stops admitting patients).

```{r}
total_time <- c()
i <- 1
while(sum(total_time) < 420){
  total_time[i] <- rexp(1,1/10)
  i <- i+1
}
total_time <- total_time[-length(total_time)] #remove the first patient that arrived after closing
cumsum(total_time) # just checking
```

The number of patients who arrived withtin opening hours this day was

```{r}
print(length(total_time))
```

We now wish to determine how many of these patients had to wait for the doctor.

```{r}
total_time
waiting_time <- c()
waiting_time[1] <- 0
rest_time <- 0
for(i in 2:length(total_time)){
  doctor_time <- runif(length(total_time), 5, 20)
  if(doctor_time[i-1]+rest_time < total_time[i]){
    waiting_time[i] <- 0
  }
  else{
    waiting_time[i] <- (rest_time+doctor_time[i-1]-total_time[i])
  }
  rest_time <- 1 
}
```




### Part b